Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Downscaling in the Swiss Alps

### 2016-01-27 Optimization case study - Journal of Hydrology/elsarticle-template.tex
Addition of the analysis on the number of analogues.

\begin{table}[htb]
	\caption{Pressure levels ($\sim$) automatically selected for the 4Zo method at the different subregions.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccccc}
			\hline ID & 300 & 400 & 500 & 600 & 700 & 850 & 925 & 1000 \\ 
			\hline 
			1  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			2  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			3  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			4  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			5  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			6  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			7  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			8  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			9  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			10 &   &   &   & $\sim$ & $\sim$ &   &   & $\sim \sim$ \\
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:levels_GA_4Zo}
\end{table}

@@

The resulting CRPSS scores are provided in Table \ref{table:scores} and are in average 35.8 \% for the CP and 35.5 \% for the VP. The improvement of the CRPSS score relatively to the reference method on the atmospheric circulation is provided in Table \ref{table:scores_diff} and is in average 15.3 \% for the CP and 9.9 \% for the VP. The score was also calculated for three precipitation thresholds: P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_4Zo}). The gain in score increases with the precipitation threshold: the relative improvement of the CRPSS is in average, respectively for the different thresholds, 13.3\%, 15.4\% and 29.1\% for the CP and 7.9\%, 11.1\% and 34.5\% for the VP. The optimization improves thus even more the forecasts for days with significant precipitation than usual days.

@@

The scores show significant improvements for both the CP and the VP (Table \ref{table:scores_diff}). Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_4Zo}) show the largest improvements. This is of particular interest in the framework of flood forecasting.

@@

\begin{table}[htb]
	\caption{Atmospheric levels automatically selected for the analogy of the atmospheric circulation ($\sim$) and the analogy of moisture ($\bullet$) of the 4Zo-2MIo method, at the different subregions.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccccc}
			\hline ID & 300 & 400 & 500 & 600 & 700 & 850 & 925 & 1000 \\ 
			\hline 
			1  & $\sim$ &   & $\sim$ &   & $\bullet \bullet$  & $\sim$ &   & $\sim$ \\
			2  & $\sim$ &   &   &   & $\sim \bullet \bullet$ & $\sim$ &   & $\sim$ \\
			3  & $\sim$ &   &   &   & $\sim \bullet \bullet$ & $\sim$ & $\sim$ &   \\
			4  &   &   & $\sim$ & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			5  &   & $\sim$ &   &   & $\sim \bullet \bullet$ &   & $\sim \sim$ &   \\
			6  &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			7  &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			8  &   &   & $\sim$ & $\bullet$ & $\sim \bullet$ &   & $\sim \sim$ &   \\
			9  &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ & $\sim$ &   \\
			10 &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:levels_GA_4Zo_2MIo}
\end{table}

@@

The CRPSS scores of the optimized methods are provided in Table \ref{table:scores} and amounts to slightly more than 40 \% for both methods and for both periods. This results in a relative improvements that ranges from 11.5 \% to 15.6 \% compared to the reference method on the moisture analogy (Table \ref{table:scores_diff}). As for 4Zo, the 4Zo-2MIo and 4Zo-4MIo methods present larger improvements in the forecast for significant rainfall (thresholds P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10). The improvement are relatively similar for 4Zo-2MIo (Table \ref{table:scores_thresholds_4Zo-2MIo}) and 4Zo-4MIo (not shown), with slightly superior scores for the latter on small precipitation (P\(\geq\)1 mm : 17.7\% and 13.0\%) and extremes (P\(\geq\)0.5\(\cdot\)P10 : 23.7\% and 29.2\%).

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. This is first visible on number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy, where the 1000~hPa that was previously systematically selected twice (Table \ref{table:levels_GA_4Zo}) is now less chosen (once or even not at all) for both 4Zo-2MIo (Table \ref{table:levels_GA_4Zo_2MIo}) and 4Zo-4MIo (not shown). There is indeed a shift of the previously selected 1000~hPa for higher levels, which is even slightly stronger with 4 moisture predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for moisture assessment, whereas it does not need it with the moisture index. This aspect has never been demonstrated before, as sequential calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

The selected pressure levels for the analogy on the moisture index are strongly centered around 700~hPa and 600~hPa. No other value has been selected when considering 2 pressure levels (Table \ref{table:levels_GA_4Zo_2MIo}), and when considering 4 levels, 850~hPa and 500~hPa are sometimes also selected (not shown). However, even in this latter method, the 700~hPa and 600~hPa levels still hold 78 \% of the selection. It is thus more efficient, in terms of forecasting performance, to consider one of this level several times at different hours rather than using another pressure level. Besides, the optimizer never chose the same pressure level at the same hour for any variable, even though it was allowed to do so.

@@

The CRPSS scores were improved by considering the moisture information (Table \ref{table:scores}). The optimized methods also perform better than the reference method based on the moisture index (Table \ref{table:scores_diff}). However, there is no drastic difference between both 4Zo-2MIo and 4Zo-4MIo methods, which suggests that considering 4 moisture predictors is not necessary. When it comes to improvements for days with precipitation above the 3 thresholds (P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the forecast, mainly for heavy rainfall.

@@

>>>>

\begin{table}[htb]
	\caption{Pressure levels ($\sim$) automatically selected for the 4Zo method at the different subregions. R represents the reference method (Table \ref{table:params_R1}).}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccccc}
			\hline ID & 300 & 400 & 500 & 600 & 700 & 850 & 925 & 1000 \\ 
			\hline 
			1  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			2  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			3  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			4  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			5  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			6  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			7  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			8  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			9  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			10 &   &   &   & $\sim$ & $\sim$ &   &   & $\sim \sim$ \\
			\hline 	
			R  &   &   & $\sim$ &   &   &   &   & $\sim$ \\
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:levels_GA_4Zo}
\end{table}

@@

The resulting CRPSS scores are provided in Table \ref{table:scores} and are in average 35.8 \% for the CP and 35.5 \% for the VP. The improvement of the CRPSS score relatively to the reference method on the atmospheric circulation (optimized by the sequential procedure) is provided in Table \ref{table:scores_diff} and is in average 15.3 \% for the CP and 9.9 \% for the VP. The score was also calculated for three precipitation thresholds: P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_4Zo}). The gain in score increases with the precipitation threshold: the relative improvement of the CRPSS is in average, respectively for the different thresholds, 13.3\%, 15.4\% and 29.1\% for the CP and 7.9\%, 11.1\% and 34.5\% for the VP. The optimization improves thus even more the forecasts for days with significant precipitation than usual days.

@@

The scores show significant improvements for both the CP and the VP (Table \ref{table:scores_diff}) compared to the reference method optimized by the sequential procedure. Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_4Zo}) show the largest improvements. This is of particular interest in the framework of flood forecasting.

@@

\begin{table}[htb]
	\caption{Atmospheric levels automatically selected for the analogy of the atmospheric circulation ($\sim$) and the analogy of moisture ($\bullet$) of the 4Zo-2MIo method, at the different subregions.  R represents the reference method (Table \ref{table:params_R2})}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccccc}
			\hline ID & 300 & 400 & 500 & 600 & 700 & 850 & 925 & 1000 \\ 
			\hline 
			1  & $\sim$ &   & $\sim$ &   & $\bullet \bullet$  & $\sim$ &   & $\sim$ \\
			2  & $\sim$ &   &   &   & $\sim \bullet \bullet$ & $\sim$ &   & $\sim$ \\
			3  & $\sim$ &   &   &   & $\sim \bullet \bullet$ & $\sim$ & $\sim$ &   \\
			4  &   &   & $\sim$ & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			5  &   & $\sim$ &   &   & $\sim \bullet \bullet$ &   & $\sim \sim$ &   \\
			6  &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			7  &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			8  &   &   & $\sim$ & $\bullet$ & $\sim \bullet$ &   & $\sim \sim$ &   \\
			9  &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ & $\sim$ &   \\
			10 &   & $\sim$ &   & $\bullet$ & $\sim \bullet$ & $\sim$ &   & $\sim$ \\
			\hline 
			R &   &   & $\sim$ &   &   & $\bullet \bullet$ &   & $\sim$ \\
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:levels_GA_4Zo_2MIo}
\end{table}

@@

The CRPSS scores of the optimized methods are provided in Table \ref{table:scores} and amounts to slightly more than 40 \% for both methods and for both periods. This results in a relative improvements that ranges from 11.5 \% to 15.6 \% compared to the reference method on the moisture analogy (Table \ref{table:scores_diff}) optimized with the sequential procedure. As for 4Zo, the 4Zo-2MIo and 4Zo-4MIo methods present larger improvements in the forecast for significant rainfall (thresholds P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10). The improvement are relatively similar for 4Zo-2MIo (Table \ref{table:scores_thresholds_4Zo-2MIo}) and 4Zo-4MIo (not shown), with slightly superior scores for the latter on small precipitation (P\(\geq\)1 mm : 17.7\% and 13.0\%) and extremes (P\(\geq\)0.5\(\cdot\)P10 : 23.7\% and 29.2\%).

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level has an influence on the parameter values of the first level. This is first visible on number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy. As for the sequential procedure, the optimal value of $N_{1}$ increases when adding a second level of analogy (Figure \ref{fig:figure_nb_analogues}). One can also notice that the optimal number of analogues $N_{2}$ for the second level of analogy of 4Zo-2HIo is slightly inferior to $N_{1}$ from 4Zo, but very close. There is a relationship between the optimal values of $N_{1}$ from both methods: the higher the $N_{1}$ of the first method, the higher $N_{1}$ and $N_{2}$ of the second is (Figure \ref{fig:figure_nb_analogues_relationships}). This relationship is not to be considered perfectly robust and should not be transposed to another case study, but in this case, it relates the magnitudes of the numbers on the various levels and methods. The optimal final number of analogues do not vary much: $23 \leq N_{1} \leq 33$ for 4Zo and $21 \leq N_{2} \leq 28$ for 4Zo-2HIo. However, the optimal number of $N_{1}$ analogues of the first level of 4Zo-2HIo varies to a greater extent: $48 \leq N_{1} \leq 84$. In this latter, it may be problematic to consider a fixed and unique value for all regions.

As for the pressure levels, the 1000~hPa that was previously systematically selected twice (Table \ref{table:levels_GA_4Zo}) is now less chosen (once or even not at all) for both 4Zo-2MIo (Table \ref{table:levels_GA_4Zo_2MIo}) and 4Zo-4MIo (not shown). There is indeed a shift of the previously selected 1000~hPa for higher levels, which is even slightly stronger with 4 moisture predictors than with 2. This change is likely due to the fact that when considering only the circulation analogy, the method tries to take into account information that can serve as proxy for moisture assessment, whereas it does not need it with the moisture index. This aspect has never been demonstrated before, as sequential calibration tools do not allow it. It can only be assessed by a global optimization technique that can tune jointly both levels of analogy.

The selected pressure levels for the analogy on the moisture index are strongly centered around 700~hPa and 600~hPa. No other value has been selected when considering 2 pressure levels (Table \ref{table:levels_GA_4Zo_2MIo}), and when considering 4 levels, 850~hPa and 500~hPa are sometimes also selected (not shown). However, even in this latter method, the 700~hPa and 600~hPa levels still hold 78 \% of the selection. It is thus more efficient, in terms of forecasting performance, to consider one of this level several times at different hours rather than using another pressure level. Besides, the optimizer never chose the same pressure level at the same hour for any variable, even though it was allowed to do so. The selected pressure levels for the analogy on the moisture index differ from the parameters resulting from the reference method optimized by the sequential procedure (Tables \ref{table:params_R2} and \ref{table:levels_GA_4Zo_2MIo}, last row). It is possible that the level 850~hPa is more optimal for the region for which the reference parameterization was established. However, this selection is usually not reconsidered when applying the sequential procedure.

@@

The CRPSS scores were improved by considering the moisture information (Table \ref{table:scores}). The optimized methods also perform significantly better than the reference method based on the moisture index (Table \ref{table:scores_diff}) and optimized by the sequential procedure. However, there is no drastic difference between both 4Zo-2MIo and 4Zo-4MIo methods, which suggests that considering 4 moisture predictors is not necessary. When it comes to improvements for days with precipitation above the 3 thresholds (P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the forecast compared to reference method, mainly for heavy rainfall.

@@

\begin{figure}[htb]
	\centerline{\includegraphics[width=\linewidth]{figures/figure_nb_analogues.pdf}}
	\caption{Number of analogues for the different regions and the various methods resulting from the optimization. Method 4Zo is made of a single level of analogy with $N_{1}$ analogues, whereas 4Zo-2HIo has two levels of analogy with respectively $N_{1}$ and $N_{2}$ analogues.}
	\label{fig:figure_nb_analogues}
\end{figure}

\begin{figure}[htb]
	\centerline{\includegraphics[width=\linewidth]{figures/figure_nb_analogues_relationships.pdf}}
	\caption{Relationships between the optimal number of analogues for both levels of the 4Zo-2HIo method and the corresponding ones for the unique level of the 4Zo method.}
	\label{fig:figure_nb_analogues_relationships}
\end{figure}
